---
title: "ShinyApp for CMPD"
author: "Paula Bailey"
date: "2022-11-27"
output: html_document
---

```{r}
library(tidyverse)
library(data.table)
library(lubridate)
library(caret)
```


```{r eval=FALSE}
shiny::runGitHub('shiny_CMPD', 'pmb-7684', ref = "main")
```


https://data.charlottenc.gov/datasets/charlotte::cmpd-incidents-1/about


Center for Data and Analytics
Powered by Esri
CMPD Incidents

Charlotte NC (IT)
City of Charlotte
Summary
For official crime statistics, please visit CMPD's Crime Statistics page at:https://charlottenc.gov/CMPD/Safety/Pages/CrimeStats.aspx . Includes all CMPD incident report types, both criminal and non-criminal. Many reports are taken only to fully document a non-criminal circumstance like a missing person, lost/missing property, etc. Other reports are only taken to document the recovery of vehicles stolen in other jurisdictions. Each incident is classified based on FBI NIBRS standards by applying a national crime hierarchy to choose the highest offense assigned to each report. More information about NIBRS standards can be found on the FBI website. Cases where Highest NIBRS Code / Highest NIBRS Description is non-criminal offense (codes in the 800 series) should not be included in analysis of total “criminal” incident reports. In addition, data includes incidents with any clearance status, including unfounded cases. A clearance status of “Unfounded” means the report has been investigated and determined either to be a false report or to involve circumstances that do not actually constitute a crime.

https://data.charlottenc.gov/datasets/charlotte::cmpd-incidents-1/about

```{r}
df <- read.csv("CMPD_Incidents11282022.csv")

```

```{r eval = FALSE}
df <- df %>% dplyr::select(-CITY, -STATE, -ZIP, -X, -Y, -X_COORD_PUBLIC, -Y_COORD_PUBLIC, -ADDRESS_DESCRIPTION, -Shape, -GlobalID, -LOCATION,  - HIGHEST_NIBRS_CODE, -OBJECTID, -CLEARANCE_DETAIL_STATUS, -LATITUDE_PUBLIC, -LONGITUDE_PUBLIC, -DATE_INCIDENT_BEGAN,-DATE_INCIDENT_END, -INCIDENT_REPORT_ID)
```

#if not OPEN, then we are going to change everyone to CLOSED. we can use mutate
```{r}
STATUS1 <- ifelse(df$CLEARANCE_STATUS %in% c('Open','OPEN'), "Open", "Closed")
```

```{r}
df2 <- mutate(df,STATUS1)
```

```{r}
df2 = df2 %>% mutate(STATUS = case_when(`STATUS1` == 'Closed'~ 0,
                                             `STATUS1` == 'Open'~ 1))
```

#if closed, how many days does it take
```{r}
DAYS_CLOSED 
```

```{r}
df2$month <- format(as.Date(df2$DATE_REPORTED, format="%Y/%m/%d"),"%m")
```


```{r}
df2 <- df2 %>% dplyr::select(-DATE_REPORTED, -CLEARANCE_DATE, - CLEARANCE_STATUS, -STATUS1)
```

```{r}
summary(df2)
```


```{r}
head(df2)
```

```{r}
sum(is.na(df2$YEAR))
sum(is.na(df2$DIVISION_ID))
sum(is.na(df2$CMPD_PATROL_DIVISION))
sum(is.na(df2$NPA))
sum(is.na(df2$LOCATION_TYPE_DESCRIPTION))
sum(is.na(df2$PLACE_TYPE_DESCRIPTION))
sum(is.na(df2$PLACE_DETAIL_DESCRIPTION))
sum(is.na(df2$STATUS))
sum(is.na(df2$HIGHEST_NIBRS_DESCRIPTION))
sum(is.na(df2$month))
```



```{r}
df2 <- df2%>%mutate(YEAR = as.factor(YEAR), DIVISION_ID = as.factor(DIVISION_ID), NPA = as.factor(NPA), CMPD_PATROL_DIVISION = as.factor(CMPD_PATROL_DIVISION), LOCATION_TYPE_DESCRIPTION = as.factor(LOCATION_TYPE_DESCRIPTION),PLACE_TYPE_DESCRIPTION = as.factor(PLACE_TYPE_DESCRIPTION), PLACE_DETAIL_DESCRIPTION = as.factor(PLACE_DETAIL_DESCRIPTION), HIGHEST_NIBRS_DESCRIPTION = as.factor(HIGHEST_NIBRS_DESCRIPTION),STATUS = as.factor(STATUS), month = as.factor(month))
```





```{r}
unique(as.character(df2$DIVISION_ID))
```

```{r}
unique(as.character(df2$CMPD_PATROL_DIVISION))
```
#Make sure to remove NA and Unknowns from the patrol division

```{r}
unique(as.character(df2$LOCATION_TYPE_DESCRIPTION))
```


```{r}
unique(as.character(df2$PLACE_TYPE_DESCRIPTION))
```


```{r}
unique(as.character(df2$PLACE_DETAIL_DESCRIPTION))
```

```{r}
as.data.frame(table(df2$PLACE_DETAIL_DESCRIPTION))
```


```{r}
unique(as.character(df2$HIGHEST_NIBRS_DESCRIPTION))
```


```{r}
as.data.frame(table(df2$HIGHEST_NIBRS_DESCRIPTION))
```

```{r}
summary(df2)
```

#*************************************************************************************************************#
# Pull info from Year 2022 

```{r}
YEAR <- 2022
yeart <- df2 %>% select(YEAR) %>% distinct() %>% pull()
df2022 <-  df2 %>% 
      filter(YEAR == yeart)
```



#remove all rows with a missing value; I don't care whihc column
```{r}
df2022 <- df2022 %>% drop_na()
```


```{r}
str(df2022)
```



```{r}
as.data.frame(table(df2022$HIGHEST_NIBRS_DESCRIPTION))
```

Using HIGHEST_NIBRS_DESCRIPTION to remove any groupings less than 2000
```{r}
df2022R <- filter(df2022, HIGHEST_NIBRS_DESCRIPTION %in% c('Aggravated Assault', 'All Other Offenses','All Other Thefts', 'Burglary/B&E','Damage/Vandalism Of Property','Drug/Narcotic Violations','Intimidation','Missing Person','Motor Vehicle Theft','Other Unlisted Non-Criminal',  'Shoplifting', 'Simple Assault', 'Theft of Motor Vehicle'))
```



```{r}
as.data.frame(table(df2022$PLACE_DETAIL_DESCRIPTION))
```

Using PLACE_DETAIL_DESCRIPTION to remove any groupings less than 2000
```{r}
df2022R <- filter(df2022R, PLACE_DETAIL_DESCRIPTION %in% c('Apartment/Duplex Private Res', 'Convenience Store','Department Store', 'Gas Station','Grocery Store/Supermarket','Hotel/Motel','Other - Commercial Place','Other - Open Area','Other - Retail','Private Residence',  'Restaurant/Diner/Coffee Shop', 'Street/Highway'))
```


```{r}
str(df2022R)
```

okay, this dataset is still very large and will take a while to run.  so let's subset a quarter of the year(either summer 7,8,9 OR fall 10,11,12)

```{r}
df2022R2 <- df2022R %>% filter(month %in% c('07','08','09'))
```

#double checking that unknown and NAs are removed
```{r}
unique(as.character(df2022R2$CMPD_PATROL_DIVISION))
```

```{r}
colnames(df2022R2) = c("YEAR", "DIVISION_ID", "DIVISION", "NPA", "LOCATION", "PLACE_TYPE", "PLACE_DETAIL","NIBRS", "STATUS", "MONTH")
```


```{r}
head(df2022R2)
```

```{r}
fwrite(df2022R2, "df2022.csv")
```



```{r}
df <- read.csv("df2022.csv")
head(df)
```
---------------------------------------------------------------------------------------------------------------
12/31/2022

```{r}
df <- read.csv("df2022.csv")
head(df)
```
interesting, the data loaded does not retain factors.
```{r}
str(df)
```



We need to get rid of NPA (not important); update YEAR, DIVISION_ID and MONTH to char

fix type
```{r}
df2 <- df%>% mutate(YEAR = as.character(YEAR), DIVISION_ID = as.character(DIVISION_ID), MONTH =as.character(MONTH))%>% dplyr::select(-NPA)
```

```{r}
df2 <- df2%>%mutate(YEAR = as.factor(YEAR), DIVISION_ID = as.factor(DIVISION_ID), DIVISION = as.factor(DIVISION), LOCATION = as.factor(LOCATION),PLACE_TYPE = as.factor(PLACE_TYPE), PLACE_DETAIL = as.factor(PLACE_DETAIL), NIBRS = as.factor(NIBRS),STATUS = as.factor(STATUS), MONTH = as.factor(MONTH))
```

```{r}
summary(df2)
```

```{r}
str(df2)
```

confirm everything is factor (above)
```{r}
  #create train and test for classification selected data

    trainIndex_C <- createDataPartition(df2$STATUS, p = .75, list = FALSE)

    train_C <- df2[trainIndex_C,]

    test_C  <- df2[-trainIndex_C,]

```
mistake to fix - use - rather than !; data is now split correctly

```{r}
# fit classification tree
  newdata <- train_C

  treefit <- train(STATUS ~ ., data = newdata, method = "rpart", 
                       preProcess = c("center", "scale"),
                       trControl = trainControl(method = "cv", number = 5))


```  

```{r}  
# classification tree summary

    plot(treefit$finalModel, main = "Classification Tree")
    text(treefit$finalModel, pretty = 0, cex = 0.6)

```

okay, so this works;  try it with OPEN and CLOSE as the response variables

```{r}
df_ <- read.csv("df2022.csv")
head(df_)
```


```{r}
df_ <- df_ %>%mutate(YEAR = as.factor(as.character(YEAR)), DIVISION_ID = as.factor(as.character(DIVISION_ID)), DIVISION = as.factor(DIVISION), LOCATION = as.factor(LOCATION),PLACE_TYPE = as.factor(PLACE_TYPE), PLACE_DETAIL = as.factor(PLACE_DETAIL), NIBRS = as.factor(NIBRS),STATUS = as.factor(STATUS), MONTH = as.factor(as.character(MONTH))) %>% dplyr::select(-NPA)
```

```{r}
str(df_)
```

confirm everything is factor (above)
```{r}
  #create train and test for classification selected data

    trainIndex_C <- createDataPartition(df_$STATUS, p = 0.75, list = FALSE)

    train_C <- df_[trainIndex_C,]

    test_C  <- df_[-trainIndex_C,]

```
mistake to fix - use - rather than !; data is now split correctly

```{r}
# fit classification tree
  newdata <- train_C

  treefit <- train(STATUS ~ ., data = newdata, method = "rpart", 
                       preProcess = c("center", "scale"),
                       trControl = trainControl(method = "cv", number = 5))


```  

```{r}  
# classification tree summary

    plot(treefit$finalModel, main = "Classification Tree")
    text(treefit$finalModel, pretty = 0, cex = 0.6)

```

############# RF

```{r}
df <- read.csv("df2022_.csv")
head(df)
```
```{r}
str(df)
```

```{r}
df <- df %>%mutate(YEAR = as.factor(as.character(YEAR)), DIVISION_ID = as.factor(as.character(DIVISION_ID)), DIVISION = as.factor(DIVISION), LOCATION = as.factor(LOCATION),PLACE_TYPE = as.factor(PLACE_TYPE), PLACE_DETAIL = as.factor(PLACE_DETAIL), NIBRS = as.factor(NIBRS),STATUS = as.factor(STATUS), MONTH = as.factor(as.character(MONTH))) %>% dplyr::select(-NPA)
```

```{r}
str(df)
```


```{r}
summary(df)
```

```{r}
  #create train and test for classification selected data

    trainIndex_R <- createDataPartition(df$STATUS, p = 0.75, list = FALSE)

    train_R <- df[trainIndex_R,]

    test_R  <- df[-trainIndex_R,]

```

```{r}
  newdata_R <- test_R
  
  
  fitrf <- train(STATUS ~ ., data = newdata_R, method = "rf", 
                 trControl = trainControl(method = "cv", number = 5),
                 tuneGrid = expand.grid(mtry = c(1:round(sqrt(ncol(newdata_R)-1)))))
```

```{r}
  fitrf2 <- randomForest(STATUS ~ ., data = newdata_R, importance = TRUE)
```

```{r}
fitrf
```


```{r}
  varimport <- varImp(fitrf)
  plot(varimport)
```



```{r}
fitrf2
```

```{r}
importance(fitrf2)
```


```{r}
varImpPlot(fitrf2)
```

using one-hot
```{r}
#define one-hot encoding function
dummy <- dummyVars(" ~ .", data=df)

#perform one-hot encoding on data frame
final_df <- data.frame(predict(dummy, newdata=df))
```

```{r}
head(final_df)
```

using one-hot
```{r}
#define one-hot encoding function
dummy <- dummyVars(" ~ .", data=df)

#perform one-hot encoding on data frame
final_df <- data.frame(predict(dummy, newdata=df))
```

```{r}
head(final_df)
```